The effect of clumpy outflows and disc anisotropy on quasar unification scenarios

### 2015-07-08 paper.tex
started writing paper proper

Outflows are ubiquitous in active galactic nuclei (AGN) and quasi-stellar objects (QSOs).
They can take the form of highly collimated radio jets \citep{bellonijet2010}, 
or mass-loaded `winds' emanating from the accretion disc. 
Approximately $20\%$ of QSOs exhibit broad absorption lines (BALs) in the ultraviolet,
providing clear evidence for outflowing absorbing material
\citep{weymann1991, knigge2008, turnermiller2009,allen2011}.
The simplest explanation for the incidence of 
BALs in quasar samples is in terms of an accretion disc wind (ADW). 
In this paradigm, the BALQSO fraction is associated with
the covering factor of the outflow.

It has also suggested that the  diverse phenemonology
of luminous AGN and QSOs can be broadly explained by
into a simple picture of geometric unification \citep[e.g.]{MCGV95, elvis2000}. 
In such a model, a biconical wind rises from 
a thin accretion disc \cite{shakurasunyaev1973}, approaching velocities
close to $\sim0.1c$. Depending on viewing angle, an observer 
may then see a BALQSO, LoBAL-QSO or normal `Type 1' quasar- with 
broad emission lines superposed on a disc-like continuum.

The UV spectrum of {\em non-BAL} QSOs is typified by a series
of strong emission lines (e.g. \la, \civ, \nv) with an underlying blue continuum
- the so-called {\sl `big blue bump'} (BBB). The BBB is normally attributed to blackbody-like
emission from an accretion disc surrounding the central black hole (REF), similar
to that described by \cite{shakurasunyaev1973}. However,
a number of issues have arisen relating to this model. First, 
AGN/QSO spectra exhibit a `break' in the spectrum at around $1000$~\AA 
which scales only weakly with black hole mass. This potentially 
suggests a problem with a thin disc model (e.g. Antonucci). 
However, it is possible that this is the result of incorrect intergalactic medium
(IGM) corrections (REF) or corresponds to the temperature 
at which a line-driven wind carries mass away from the system (laor davis).
Second, results from microlensing (REFs) imply that the disc emission 
region is $\sim4$ times larger than one might expect from a Shakura-Sunyaev
model. Inhomogenous discs have been proposed as an alternative (REFs).
These observations appear to pose problems for the thin disc model of luminous,
sub-Eddington AGN. However, recent results from \cite{capellupo2015} find 
that if one includes a combination of mass-loss, general relativity (GR) and Comptonisation
then AGN SEDs can, in general, be fitted well with accretion disc models.
Uncovering the intrinsic disc SED and understanding the effect of the outflow on the accretion 
mechanism is clearly crucial if we are to properly understand the physical nature
of AGN.

The geometry and size of the BLR is also a matter of contention. 
The main constraints on the emission region size come from microlensing (REFs)
and reverberation mapping (REFs). While these observations have mostly been
carried out for Seyfert galaxies, there are also a few instances of these methods
being applied to quasars (REFs). A number of different proposals have been made 
for the origin of the BLR. Early AGN unification scenarios posited that the broad emission lines
where produced by clouds of plasma orbiting fairly close to the BH (refs).
Since then, multiple interpretations of a disc wind model have been proposed,
with varying radii and geometries (refs).

>>>>

\clearpage
Quasars and luminous active galactic nuclei (AGN) 
exhibit spectral energy distributions (SEDs) that
typically exhibit a series of strong emission lines (e.g. \la, \civ, \nv) 
with an underlying blue continuum - the so-called {\sl `big blue bump'} (BBB). 
The BBB is normally attributed to 
emission from a geometrically thin, optically thick accretion disc 
surrounding the central black hole (REF), similar
to that described by \cite{shakurasunyaev1973}.

In addition to the inflowing accreting material, 
outflowing material is ubiquitous in active galactic nuclei (AGN) 
and quasi-stellar objects (QSOs). These outflows can take the form of 
highly collimated radio jets \citep{bellonijet2010}, 
or mass-loaded `winds' emanating from the accretion disc. 
Approximately $20\%$ of QSOs exhibit broad absorption lines (BALs) in the ultraviolet,
providing clear evidence for outflowing absorbing material
\citep{weymann1991, knigge2008, turnermiller2009, allen2011}.
The simplest explanation for the incidence of 
BALs in quasar samples is in terms of an accretion disc wind (ADW). 
Within this paradigm, the BALQSO fraction is associated with
the covering factor of the outflow.

It is possible that the  diverse phenemonology
of luminous AGN and QSOs can be broadly explained by
into a simple picture of geometric unification \citep[e.g.][]{MCGV95, elvis2000}. 
In such a model, a biconical wind rises from 
the accretion disc, and the class of object is explained by the material
intercepting the line of sight. Depending on viewing angle, an observer 
may then see a BALQSO, LoBAL-QSO or normal `Type 1' quasar.

As well as acting as a source of photoionized plasma, 
a disc wind may also have a profound effect on the structure and 
emergent continuum of the accretion disc itself.
Mass-loss will alter the accretion rate and resultant 
temperature of the accretion disc, possibly explaining some 
of the features we typically see in luminous AGN (REF).
Recent results from \cite{capellupo2015} find 
that if one includes a combination of mass-loss, general relativity (GR) and Comptonisation
then AGN SEDs can, in general, be fitted well with accretion disc models.
In general, mass-loss treatments appear to be critical if an $\alpha$-disc
model is to successfully fit AGN SEDs, particular in the UV region of the spectrum.

% The UV spectrum of {\em non-BAL} QSOs is typified by a series
% of strong emission lines (e.g. \la, \civ, \nv) with an underlying blue continuum
% - the so-called {\sl `big blue bump'} (BBB). The BBB is normally attributed to blackbody-like
% emission from an accretion disc surrounding the central black hole (REF), similar
% to that described by \cite{shakurasunyaev1973}. However,
% a number of issues have arisen relating to this model. First, 
% AGN/QSO spectra exhibit a `break' in the spectrum at around $1000$~\AA 
% which scales only weakly with black hole mass. This potentially 
% suggests a problem with a thin disc model (e.g. Antonucci). 
% However, it is possible that this is the result of incorrect intergalactic medium
% (IGM) corrections (REF) or corresponds to the temperature 
% at which a line-driven wind carries mass away from the system (laor davis).
% Second, results from microlensing (REFs) imply that the disc emission 
% region is $\sim4$ times larger than one might expect from a Shakura-Sunyaev
% model. Inhomogenous discs have been proposed as an alternative (REFs).
% These observations appear to pose problems for the thin disc model of luminous,
% sub-Eddington AGN. However, recent results from \cite{capellupo2015} find 
% that if one includes a combination of mass-loss, general relativity (GR) and Comptonisation
% then AGN SEDs can, in general, be fitted well with accretion disc models.
% Uncovering the intrinsic disc SED and understanding the effect of the outflow on the accretion 
% mechanism is clearly crucial if we are to properly understand the physical nature
% of AGN.

% The geometry and size of the BLR is also a matter of contention. 
% The main constraints on the emission region size come from microlensing (REFs)
% and reverberation mapping (REFs). While these observations have mostly been
% carried out for Seyfert galaxies, there are also a few instances of these methods
% being applied to quasars (REFs). A number of different proposals have been made 
% for the origin of the BLR. Early AGN unification scenarios posited that the broad emission lines
% where produced by clouds of plasma orbiting fairly close to the BH (refs).
% Since then, multiple interpretations of a disc wind model have been proposed,
% with varying radii and geometries (refs).
